---
title: "awLab03"
author: "Aisha"
date: "September 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
data(happy, package = 'productplots')
head(happy)

HAPPY <- readRDS("HAPPY.rds")
head(HAPPY)
```

# Question 1
```{r}
HAPPY$POLVIEWS[(HAPPY$POLVIEWS %in% c('IAP','DK','NA'))] = NA
```

# Question 2 
```{r}
summary(HAPPY)

HAPPY$HAPPY <- as.factor(HAPPY$HAPPY)
HAPPY$SEX <- as.factor(HAPPY$SEX)
HAPPY$MARITAL <- as.factor(HAPPY$MARITAL)
HAPPY$DEGREE <- as.factor(HAPPY$DEGREE)
HAPPY$FINRELA <- as.factor(HAPPY$FINRELA)
HAPPY$HEALTH <- as.factor(HAPPY$HEALTH)
HAPPY$PARTYID <- as.factor(HAPPY$PARTYID)
HAPPY$POLVIEWS <- as.factor(HAPPY$POLVIEWS)

HAPPY$AGE[HAPPY$AGE == '89 OR OLDER'] <- '89'
HAPPY$AGE <- as.numeric(HAPPY$AGE)

summary(HAPPY)
```

# Question 3
```{r}
HAPPY$HAPPY <- reorder(HAPPY$HAPPY, HAPPY$AGE, mean)
HAPPY$SEX <- reorder(HAPPY$SEX, HAPPY$AGE, mean)
HAPPY$MARITAL <- reorder(HAPPY$MARITAL, HAPPY$AGE, mean)
HAPPY$DEGREE <- factor(HAPPY$DEGREE,levels(HAPPY$DEGREE)[c(6,4,5,1,3,2)])
HAPPY$FINRELA <- reorder(HAPPY$FINRELA, HAPPY$AGE, mean)
HAPPY$HEALTH <- factor(HAPPY$HEALTH,levels(HAPPY$HEALTH)[c(6,3,4,2,1,5)])
HAPPY$PARTYID <- reorder(HAPPY$PARTYID, HAPPY$AGE, mean)
HAPPY$POLVIEWS <- reorder(HAPPY$POLVIEWS, HAPPY$AGE, mean)
```

# Aisha's Ivestigation  
How are Health, Marital Status, and Happiness related?
```{r}
library(ggplot2) #AGE and HEALTH

ggplot(HAPPY, aes(x = HEALTH)) +
  geom_bar(aes(fill = MARITAL)) +
  facet_wrap(~HAPPY) +
  coord_flip()
```

Overall we can see that most people are "Pretty Happy" and less people are "Not Too Happy". In the Happy categories, people are mainly "Married" (or "Never Married") with most people in "Good" or "Excellent" health.  Meanwhile the "Not Too Happy" category doesn't appear to have any trends (though there are possibly more people in "Good" or "Fair"  health than in the other categories).